---
title: "ST558 Porject 1: Part 3"
author: "Bridget Knapp and Chien-Lan Hsueh"
institute: Online Statistics Master Program, NCSU
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: cerulean
    highlight: haddock
    code_folding: none
  html_document:
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.path = "./images/")
```


## Summarizations

> You should produce some basic (but meaningful) summary statistics and plots about the training data you
are working with (especially as it relates to your response).
>
> As you will automate this same analysis across other data, you can’t describe the trends you see in the graph (unless you want to try to automate that!). You should describe what to look for in the summary statistics/plots to help the reader understand the summary or graph. Ex: A scatter plot with the number of shares on the y-axis and the positive word rate on the x-axis is created:
>
> 'We can inspect the trend of shares as a function of the positive word rate. If the points show an upward trend, then articles with more positive words tend to be shared more often. If we see a negative trend then articles with more positive words tend to be shared less often.'
>
> Each group member is responsible for producing some summary statistics (means, sds, contingency tables, etc.) and for producing at least three graphs (each) of the data.


"part 2 - data.Rmd"


```{r, include = FALSE}
knitr::knit("part 2 - data.Rmd", output = tempfile())
```



## Start here...

```{r}
# check balance - compare ecdf of train and test sets
plot(ecdf(df_train$shares), col = "blue", main = "Empirical Cumulative Distribution", xlab = "shares")
plot(ecdf(df_test$shares), col = "red", add = T)
```


```{r}
table(df_train$channel)
channel <- "lifestyle"          

# subset by channel
df_train_x <- df_train %>% filter(channel == channel)
df_test_x <- df_test %>% filter(channel == channel)
```

```{r 8:29_7_7_2022}
#A summary table showing the sample size (N), mean, standard deviation (Std. Dev.), minimum (Min), 25th percentile, 75th percentile, and maximum (Max)
library(vtable)
st(df_train_x)
```

```{r 7:31_7_7_2022}
A <- df_train_x %>%
  select(kw_avg_max,shares)

g <- ggplot(A, aes(x = kw_avg_max))
g + geom_histogram(color = "blue", fill = "red",
size = 2, binwidth = 30) +
  labs(x = "Best Keyword (Average Number of Shares)", title = "Distribution of Shares by Best Keyword", y="Number of Shares")
```

```{r 8:10_7_7_2022}
#we need to be able to cycle through the channels
#maybe something like this: A <- df_train_x %>% filter(channel==channel)
A <- df_train_x %>% filter(channel=="lifestyle") 

A <- dplyr::select_if(A, is.numeric)

r <- cor(A, use="complete.obs")
library(ggplot2)
library(ggcorrplot)
ggcorrplot(r, 
            hc.order = TRUE, 
            type = "lower",
            tl.cex = 6)
```

```{r 5:40_7_8_2022}

avg_positive_polarity_Sum <- df_train_x %>% group_by(avg_positive_polarity) %>% summarize(prop_avg_positive_polarity = mean(shares),n = n())
avg_positive_polarity_Sum

ggplot(avg_positive_polarity_Sum, aes(x=avg_positive_polarity,y=prop_avg_positive_polarity,size=n)) + geom_point(stat="identity") +
  labs(x = "Average Polarity of Positive words", title = "Average Polarity of Positive words",y="Proportion of Average Polarity of Positive words")

```

```{r 5:58_7_8_2022}

g <- ggplot(df_train_x, aes(x = shares, y = self_reference_avg_sharess,col=weekday))
g + geom_point() +
  labs(x = "Number of Shares", title = "Number of Shares vs. Avg. Number of Shares of Referenced Articles in Mashable", y="Avg. Number of Shares of Referenced Articles in Mashable")


```

```{r 6:06_7_8_2022}
g <- ggplot(df_train_x, aes(x = shares, y = n_tokens_content,group=n_tokens_title,fill=shares))
g + geom_boxplot(fill = "grey") +
  labs(x = "Number of Shares", title = "Distribution of Shares Grouped by Number of Words in Article", y="Number of Words in the Article")
```





```{r}
# quick summaries of numeric and categorical variables
skim(df_train_x)
```



```{r}
#mean(df_train_x$shares)
```


Form `r channel` news channel subset, we found the mean of `shares` is `r mean(df_train_x, na.rm)`.