---
title: 'ST558 Porject 1: Part 3'
author: "Bridget Knapp and Chien-Lan Hsueh"
institute: Online Statistics Master Program, NCSU
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: cerulean
    highlight: haddock
    code_folding: none
params:
  channel: "lifestyle"
  load_data: TRUE
---


```{r, include=FALSE}
if(params$load_data){
  knitr::knit(here::here("_Rmd", "part 2 - data.Rmd"), output = tempfile())
  
  df_train_x <- df_train %>% filter(channel == params$channel)
  df_test_x <- df_test %>% filter(channel == params$channel)
}
```



## News Channel: ``r params$channel``

In this section, a quick EDA will be done on the **training set** for news channel ``r params$channel``:

- Overlook of the data set
- Response variable
- Categorical Predictors
- Numeric Predictors


### Overlook of the data set

Verify the data sets only has data from ``r params$channel`` channel:
```{r}
table(df_train_x$channel, df_train_x$weekday)
table(df_test_x$channel, df_test_x$weekday)
```

The tables below give quick summaries of all the numeric and categorical variables:
```{r}
# quick summaries of numeric and categorical variables
skim(df_train_x)
```

### Response Variable

Next, let's learn more about our response variable: `shares` with a 5-number summary. With a histogram, we can visually see it's distribution and determine if it is symmetric, skewed left, or skewed right.

```{r histogram}
# 5-number summary on shares
summary(df_train_x$shares)

# histogram on shares
df_train_x %>% 
  ggplot(aes(x = shares)) + 
  geom_histogram(binwidth = 1000) +
  coord_cartesian(xlim = c(0, 100000)) +
  labs(
    title = paste0("Histogram: Distribution of Shares for News Channel ", params$channel),
    x="Shares",
    y="Frequency")
```

### Categorical Predictors

Despite the `channel` variable, there are two categorical variables `weekday`, `is_weekend`. A one-way contigency table shows how many articles were published on each weekday.

```{r}
df_train_x %>% 
  select(weekday, is_weekend) %>% 
  summary()
```
The following side-by-side box plots help us visualize the distributions of `shares` on each weekday.

```{r boxplot}
df_train_x %>% 
  ggplot(aes(weekday, shares)) +
  geom_boxplot() +
  scale_y_log10() +
  geom_jitter(width = 0.05) +
  labs(
    title = paste0("Boxplots: Distribution of Shares in News Channel ", params$channel, " for Each Weekday"),
    x = "Weekday")
```

### Numeric Predictors

We have `r ncol(df_train_x %>% select_if(is.numeric))` numeric variables. To further investigate the relationship among these numeric variables, we check their correlation by making pair-wise correlation plots. 

```{r cor, warning=FALSE, message=FALSE}
ggcorrplot(
  cor(select_if(df_train_x, is.numeric), use = "complete.obs"), 
  hc.order = TRUE, 
  type = "lower",
  tl.cex = 6,
  title = paste0("Correlation Plot for News Channel ", params$channel))
```

Below, we use scatter plots to inspect the relationship between the response variable `shares` and various numeric variables in related groups:

- `shares` vs. token-related variables
- `shares` vs. numbers of links, keyword, images, videos
- `shares` vs. keyword-related metrics
- `shares` vs. tone polarity metrics
- `shares` vs. title-related metrics

#### `shares` vs. Token-related Variables
```{r scatter1}
# scatter plots: shares vs. token-related variables
df_train_x %>% 
  select(shares, starts_with("n_"), contains("token")) %>% 
  pivot_longer(
    cols = -c(shares),
    names_to = "token_metric",
    values_to = "number_of_tokens") %>% 
  ggplot(aes(number_of_tokens, shares)) +
  geom_point() +
  facet_wrap(vars(token_metric), scales = "free") +
  ggtitle("Shares vs. Token-related Variables")
```

#### `shares` vs. numbers of links, keywords, images and videos
```{r scatter2}
# scatter plots: shares vs. numbers of links, keywords, images and videos
df_train_x %>% 
  select(shares, starts_with("num_"), starts_with("self_reference")) %>% 
  pivot_longer(
    cols = -c(shares),
    names_to = "number_metric",
    values_to = "number_of_links_keywords_images_videos") %>% 
  ggplot(aes(number_of_links_keywords_images_videos, shares)) +
  geom_point() +
  facet_wrap(vars(number_metric), scales = "free") +
  ggtitle("Shares vs. Numbers of Links, Keywords, Images and Videos")
```


#### `shares` vs. Keyword-related Metrics
```{r scatter3}
# scatter plots: shares vs. numbers of links, keywords, images and videos
df_train_x %>% 
  select(shares, starts_with("kw_")) %>% 
  pivot_longer(
    cols = -c(shares),
    names_to = "number_metric",
    values_to = "keyword_related_metric") %>% 
  ggplot(aes(keyword_related_metric, shares)) +
  geom_point() +
  facet_wrap(vars(number_metric), scales = "free") +
  ggtitle("Shares vs. Keyword-related Metrics")
```

#### `shares` vs. Word Polarity Metrics
```{r scatter4}
# scatter plots: shares vs. numbers of links, keywords, images and videos
df_train_x %>% 
  select(
    shares, 
    contains("subjectivity"), 
    contains("polarity"), 
    contains("positive"), 
    contains("negative"),
    -contains("title")) %>% 
  pivot_longer(
    cols = -c(shares),
    names_to = "number_metric",
    values_to = "tone_polarity") %>% 
  ggplot(aes(tone_polarity, shares)) +
  geom_point() +
  facet_wrap(vars(number_metric), scales = "free") +
  ggtitle("Shares vs. Word Polarity Metrics")
```

#### `shares` vs. Title-related Metrics
```{r scatter5}
# scatter plots: shares vs. numbers of links, keywords, images and videos
df_train_x %>% 
  select(shares, contains("title")) %>% 
  pivot_longer(
    cols = -c(shares),
    names_to = "number_metric",
    values_to = "title_related_metric") %>% 
  ggplot(aes(title_related_metric, shares)) +
  geom_point() +
  facet_wrap(vars(number_metric), scales = "free") +
  ggtitle("Shares vs. Title-related Metrics")
```