---
title: "ST558 Porject 1: Part 1"
author: "Bridget Knapp and Chien-Lan Hsueh"
institute: Online Statistics Master Program, NCSU
date: "`r Sys.Date()`"
output:
  html_notebook:
    theme: cerulean
    highlight: haddock
    code_folding: none
  html_document:
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.path = "./images/")
```


## Introduction Section

> You should have an introduction section that briefly describes the data and the variables you have to work with (just discuss the ones you want to use). Your target variables is the shares variable. You should also mention the purpose of your analysis and the methods you’ll use to model the response. You’ll describe those in more detail later. 
>
> This section should be done by the ‘second’ group member. 



## Start here...

```{r render, eval = FALSE}
#this is the code I used to render this document before I committed changes to the repository
rmarkdown::render("part 1 - introduction.Rmd",
                  output_format = "github_document",
                  output_options = list(
                    toc = TRUE,
                    html_preview=FALSE) 
)

```

# About the Data

We are going to get our data from the UCI Machine Learning Repository. Essentially, it’s a website with over 600 data sets for people to practice coding and machine learning. We are going to use the Online News Popularity Data Set, which summarizes a variety of aspects about articles published by Mashable, an American news website, over the course of two years. Some of the aspects include:

+ Number of words in the title and in the content
+ Number of links, images, and videos
+ Type, or channel, of news (Lifestyle, Entertainment, Business, Social Media, Tech, or World)
+ Minimum, maximum, and average number of shares for the worst, best, and average keyword 
+ When the article was published (Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday, or a weekend)?
+ Minimum, maximum, and average polarity of positive and negative words in the article

We will ignore the URL of the article and the number of days between the article publication and the dataset acquisition because they are non-predictive.

# What We Are Going to Do

The purpose of this repository is to automatically generate reports for each channel of news that summarize the data and to predict the number of shares based on chosen aspects of interest. To make our predictions, we will split the data up into a training and a test set and, using the training set, make a linear regression model and an ensemble tree-based model, specifically a random forest model and a boosted tree model.
